---
title: "Skills & Tools"
---

## Technical Skills Matrix

My skillset combines archaeological expertise with modern data science capabilities:

```{r}
#| echo: false
#| warning: false
#| message: false
#| fig-height: 6
#| fig-width: 10

library(tidyverse)
library(ggplot2)

# Create skills data
skills_data <- tibble(
  category = rep(c("Programming", "Geospatial", "Data Management", 
                   "Research Methods", "Communication"), each = 3),
  skill = c(
    "R", "Python", "SQL",
    "QGIS", "ArcGIS", "PostGIS",
    "PostgreSQL", "Git/GitHub", "Metadata Standards",
    "Statistical Analysis", "GIS Analysis", "Database Design",
    "Technical Writing", "Training Delivery", "Project Management"
  ),
  proficiency = c(
    95, 80, 90,
    90, 85, 85,
    85, 90, 90,
    90, 95, 85,
    95, 95, 90
  )
)

# Create visualization
ggplot(skills_data, aes(x = reorder(skill, proficiency), y = proficiency, fill = category)) +
  geom_col(show.legend = TRUE) +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(face = "bold", size = 14),
    axis.title = element_text(face = "bold"),
    legend.position = "bottom"
  ) +
  labs(
    title = "Technical Proficiency Overview",
    x = NULL,
    y = "Proficiency Level",
    fill = "Category"
  )
```

---

## Core Competencies

### {{< fa brands r-project >}} R Programming

**Expertise Level**: Advanced

**Applications**:

- Statistical analysis and modeling
- Data visualization with ggplot2
- Spatial analysis with sf, terra, raster
- Reproducible reporting with R Markdown and Quarto
- Package development
- Automation and scripting

**Example Projects**:

- Archaeological survey data analysis
- Spatial statistics for settlement patterns
- Automated report generation
- Data visualization dashboards

### {{< fa brands python >}} Python

**Expertise Level**: Intermediate-Advanced

**Applications**:

- Data processing and cleaning
- Geospatial analysis (geopandas, shapely)
- Web scraping and automation
- Database interactions
- Jupyter notebooks for analysis

### {{< fa database >}} Database Management

**PostgreSQL + PostGIS**:

- Database design and architecture
- Complex queries and data manipulation
- Spatial database operations
- Performance optimization
- Data integration pipelines

**Applications**:

- Linear Earthworks database (Durham project)
- Archaeological site databases
- Heritage data management systems

### {{< fa map >}} Geospatial Analysis

**QGIS & ArcGIS**:

- Spatial analysis and modeling
- Cartography and map production
- Least Cost Path analysis
- Viewshed and visibility analysis
- Network analysis

**PostGIS**:

- Spatial queries and operations
- Coordinate system transformations
- Geometric calculations
- Integration with PostgreSQL

### {{< fa code-branch >}} Version Control & Collaboration

**Git & GitHub**:

- Version control workflows
- Collaborative development
- Documentation with GitHub Pages
- Issue tracking and project management

---

## Research Methods

### Spatial Analysis

```{r}
#| echo: true
#| eval: false

# Example: Least Cost Path Analysis in R
library(sf)
library(terra)
library(gdistance)

# Load elevation data
dem <- rast("elevation.tif")

# Create cost surface
slope <- terrain(dem, "slope")
cost_surface <- 1 / (1 + slope)  # Inverse relationship

# Calculate transition matrix
trans <- transition(cost_surface, transitionFunction = mean, directions = 8)
trans_corrected <- geoCorrection(trans)

# Find least cost path between points
start <- c(longitude1, latitude1)
end <- c(longitude2, latitude2)
lcp <- shortestPath(trans_corrected, start, end, output = "SpatialLines")
```

### Statistical Analysis

- Exploratory data analysis
- Hypothesis testing
- Regression modeling
- Spatial statistics
- Bayesian analysis
- Time series analysis

### Data Visualization

```{r}
#| echo: true
#| eval: false

# Example: Archaeological site distribution map
library(tidyverse)
library(sf)

sites %>%
  ggplot() +
  geom_sf(aes(color = period, size = importance)) +
  scale_color_viridis_d() +
  theme_minimal() +
  labs(
    title = "Distribution of Archaeological Sites",
    subtitle = "Iron Age and Roman Britain",
    color = "Period",
    size = "Site Importance"
  )
```

---

## Tools & Technologies

### Development & Documentation

::: {.grid}

::: {.g-col-12 .g-col-md-6}
**Development**
- RStudio
- VS Code
- Jupyter Lab
- Command Line/Terminal
:::

::: {.g-col-12 .g-col-md-6}
**Documentation**
- Quarto
- R Markdown
- MkDocs
- Jupyter Notebooks
- LaTeX
:::

:::

### Data & Research Management

::: {.grid}

::: {.g-col-12 .g-col-md-6}
**Data Processing**
- tidyverse (R)
- pandas (Python)
- OpenRefine
- Excel/LibreOffice
:::

::: {.g-col-12 .g-col-md-6}
**Research Management**
- Git/GitHub
- Zotero
- OSF (Open Science Framework)
- OASIS database
:::

:::

### Specialized Heritage Tools

- Digital recording equipment (GNSS, Total Station)
- 3D modeling (Structure from Motion)
- Laser scanning
- Archaeological databases
- Heritage data standards

---

## Certifications & Training

- **Carpentries Certified Instructor**
- **Research Data Management** specialist training
- **FAIR Principles** implementation
- **Open Science** practices
- **Project Management** for research

---

## Continuous Learning

I'm committed to staying current with emerging technologies:

**Currently Exploring**:

- Advanced R package development
- Machine learning applications in archaeology
- Cloud computing for research (AWS, Google Cloud)
- Interactive web applications (Shiny, Dash)
- Linked Open Data for heritage

---

## Consulting Services

I offer consulting in:

::: {.callout-note icon=false}
## Research Data Management
- FAIR data implementation
- Metadata schema design
- Data management planning
- Repository selection and setup
:::

::: {.callout-note icon=false}
## Technical Implementation
- Database design and development
- GIS analysis and modeling
- R/Python workflow development
- Automation and scripting
:::

::: {.callout-note icon=false}
## Training & Capacity Building
- Workshop design and delivery
- Curriculum development
- Technical documentation
- Team skill assessment
:::

[{{< fa envelope >}} Get in touch](mailto:nickyjgarland@gmail.com) to discuss your project needs.
